Meta-analytic approaches and effect sizes to account for 'nuisance heterogeneity' in comparative physiology

### 2021-09-17 MS/ms.Rmd
Fixed up legend, and some description of the lit search.

We included comparative studies / meta-analyses that were focused on physiological questions. Narrative reviews and comments were not considered because they did not collate quantitative data from the literature. Studies needed to have collected physiological trait data, such as hormone levels, thermal physiological parameters (Critical thermal maximum (CTmax), critical thermal minimum (CTmin), thermal preferences, thermal breadth, lethal thermal tolerance limit (LT50), lethal thresholds), metabolic rates (routine, maximum, and minimum metabolic rate, aerobic scope), enzyme reaction rates and concentrations, oxidative stress measures and consequences (e.g., DNA damage, teleomere attrition), performance measurements (e.g., heart rates, sprint speed, swim speed, growth rates), body composition (e.g., fatty acid profiles) and immune traits (e.g., cell function, antibodies). Studies were also included if they measured the relationship between physiological traits and body size (e.g., metabolic scaling), reproductive allocation, or survival across taxa. Studies on behavior were only included if the synthesis was explicitly interested in the proximate physiological drivers of behavioral variation. We also included meta-analyses that quantified plasticity in physiological traits in response to environmental variation. While many studies in comparative physiology collect data as part of a comparative analysis, we focus here only on studies collating data from existing literature because this is the major data source in quantitative syntheses. We excluded quantitative syntheses interested in phenology, bio-mechanics, sexual size dimorphism (i.e., except if the study was interested in sexual differences in physiological traits), those focused on validating physiological methodology, or characterising population genetic structure, gene transcription, and quantitative genetics. There were also meta-analyses that focused on colour evolution which varied in their goals but were most often not focused on physiological questions. As such, these studies were also excluded. We restricted our focus to animal studies because most plant synthesis papers were focused on ecological (e.g., quantified changes in abundance and distribution of species) or agricultural questions (e.g., studies on soil microbial function, nitrogen and phosphorus capture, methane output, and greenhouse gas emissions).

@@

```{r, figsurv, fig.wid = 16.833333, fig.height=5,fig.cap="**Summary of meta-analytic practices in comparative physiology.** (A) Different types of effect measures commonly used in meta-analyses and their relative frequency of use. (B) (B) Frequencies of studies where ‘nuisance heterogenity’ had the potential to impact effect sizes (x-axis), and the frequencies of studies explicitly dealing with nuisance heterogeneity (bar colour). (C) Frequencies of analytical approaches used to analyse effect measures. Note that only ‘Weighted Regression’, ‘Random Effects Models’, ‘Multilevel Metaregression Models’ and ‘Multilevel Meta-analysis Models’ are weighted models. The results of this survey were drawn from 62 studies in comparative physiology."}
p1 = dat %>% dplyr::group_by(effect_size_type) %>% janitor::tabyl(effect_size_type)  %>% 
  dplyr::summarise(SMD = dplyr::if_else(grepl("SMD", effect_size_type), n, 0),
                  lnRR = dplyr::if_else(grepl("Response Ratio", effect_size_type), n, 0),
                    Zr = dplyr::if_else(grepl("correlation", effect_size_type), n, 0),
                 Other = dplyr::if_else(grepl("other|Q10", effect_size_type), n, 0),
                 Mean = dplyr::if_else(grepl("mean/proportion/raw", effect_size_type), n, 0),
                 lnCVR = dplyr::if_else(grepl("lnCVR", effect_size_type), n, 0),
                 Slope = dplyr::if_else(grepl("slope", effect_size_type), n, 0)) %>% 
  colSums(.) %>% data.frame(effect = names(.), n = .) %>% mutate(prop = n / sum(n)) %>% arrange(effect) %>% 
  mutate(effect = fct_relevel(effect, c("Other", "Zr", "lnRR", "SMD", "lnCVR", "Slope", "Means"))) %>% 
  ggplot() + tm + 
  geom_col(aes(x = effect, y = prop), width = 0.5, fill = "black", colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,0.35)) +
  labs(x = "Effect Size",
       y = "Frequency") +
  coord_flip()

>>>>

We included comparative studies / meta-analyses that were focused on physiological questions. Narrative reviews and comments were not considered because they did not collate quantitative data from the literature. Studies needed to have collected physiological trait data, such as hormone levels, thermal physiological parameters (Critical thermal maximum (CTmax), critical thermal minimum (CTmin), thermal preferences, thermal breadth, lethal thermal tolerance limit (LT50), lethal thresholds), metabolic rates (routine, maximum, and minimum metabolic rate, aerobic scope), enzyme reaction rates and concentrations, oxidative stress measures and consequences (e.g., DNA damage, teleomere attrition), performance (e.g., maximum heart rates, sprint speed, swim speed) and growth rates, body composition (e.g., fatty acid profiles) and immune traits (e.g., cell function, antibodies). Studies were also included if they measured the relationship between physiological traits and body size (e.g., metabolic scaling), reproductive allocation, or survival across taxa. Studies on behavior were only included if the synthesis was explicitly interested in the proximate physiological drivers of behavioral variation. We also included meta-analyses that quantified the change in physiological traits in response to environmental variation (i.e., phenotypic plasticity). While many studies in comparative physiology collect data as part of a comparative analysis, we focus here only on studies collating data from existing literature because this is the major data source in quantitative syntheses. We excluded quantitative syntheses interested in phenology, biomechanics, sexual size dimorphism (i.e., except if the study was interested in sexual differences in physiological traits), those focused on validating physiological methodology, or characterising population genetic structure, gene transcription, and quantitative genetics. There were also meta-analyses that focused on colour evolution which varied in their goals but were most often not focused on physiological questions. As such, these studies were also excluded. We restricted our focus to animal studies because most plant synthesis papers were focused on ecological (e.g., quantified changes in abundance and distribution of species) or agricultural questions (e.g., studies on soil microbial function, nitrogen and phosphorus capture, methane output, and greenhouse gas emissions).

@@

```{r, figsurv, fig.wid = 16.833333, fig.height=5,fig.cap="**Summary of meta-analytic practices in comparative physiology.** (A) Different types of effect measures commonly used in meta-analyses and their relative frequency of use. (B) Frequency of studies where ‘nuisance heterogenity’ had the potential to impact effect sizes (x-axis), and the frequencies of studies explicitly dealing with nuisance heterogeneity (bar colour). (C) Frequencies of analytical approaches used to analyse effect measures. Note that only ‘Weighted Regression’, ‘Random Effects Models’, ‘Multilevel Metaregression Models’ and ‘Multilevel Meta-analysis Models’ are weighted models. The results of this survey were drawn from 62 studies in comparative physiology."}
p1 = dat %>% dplyr::group_by(effect_size_type) %>% janitor::tabyl(effect_size_type)  %>% 
  dplyr::summarise(SMD = dplyr::if_else(grepl("SMD", effect_size_type), n, 0),
                  lnRR = dplyr::if_else(grepl("Response Ratio", effect_size_type), n, 0),
                    Zr = dplyr::if_else(grepl("correlation", effect_size_type), n, 0),
                 Other = dplyr::if_else(grepl("other|Q10", effect_size_type), n, 0),
                 Mean = dplyr::if_else(grepl("mean/proportion/raw", effect_size_type), n, 0),
                 lnCVR = dplyr::if_else(grepl("lnCVR", effect_size_type), n, 0),
                 Slope = dplyr::if_else(grepl("slope", effect_size_type), n, 0)) %>% 
  colSums(.) %>% data.frame(effect = names(.), n = .) %>% mutate(prop = n / sum(n)) %>% arrange(effect) %>% 
  mutate(effect = fct_relevel(effect, c("Other", "Zr", "lnRR", "SMD", "lnCVR", "Slope", "Means"))) %>% 
  ggplot() + tm + 
  geom_col(aes(x = effect, y = prop), width = 0.5, fill = "black", colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,0.35)) +
  labs(x = "Effect Size",
       y = "Frequency") +
  coord_flip()
